Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-01-12 paper/paper.tex
pass

\tname{} uses a composition of heurisitc functions to sort the rules
it produces. Each function uses a different criterion to estimate how
similar a rule, created in the previous step, is to rules from a
public database~\cite{emerging-threats-open} including a total of
\numrulessuri\ rules. We describe in the following each of these
heuristic functions. Each function $f_i$ ranges in the 0-1
domain. \tname{} uses the weighted average of these functions to sort
rules: $\sum{f_i}/|f_i|$.

\subsubsection{Size of a rule}
One element that \tname{} uses to measure similarity of synthesized to
human-created rules is the size of the synthetic rule. Informally,
rules with popular sizes score high whereas inpopular rules score low.

Figure~\ref{fig:distribution-rule-size} shows the histogram of rule
sizes for our rule dataset. The x-axis shows the number of options
included in a rule (up to 10) while the y-axis shows the number of
rules with the corresponding number of options. Note that there are 11
rules with no options. These rules only use the header to signal a
traffic as malicious. Also note that most rules have 2 to 6
options. \tname{} values rules in that size range higher compared to
other sizes.

@@

\noindent
, where $\mathit{len(r)}$ denotes the number of options of a rule $r$,
$\mathit{rules\_per\_size}$ encodes distribution of sizes as a
dictionary mapping sizes to number of associated rules, and
$\mathit{max(rules\_per\_size)}$ returns the maximum number of rules
for any dictionary entry. Consequently, the contribution of size is
proportional to its popularity. For example, the contribution of this
heuristic for a rule with 3 options would be 0.684 (=3,662/5,353)
whereas the contribution of a rule with 9 options would be 14.25
(=763/5,353). Note that the measurement is normalized in the 0-1
range.

\subsubsection{Number of content options.} We empirically observed that content options are very
common. As consequence, we defined a heuristic function that takes the
number of content options into account. We calculate the contribution
of number of content options with the following formula:

@@

\[\frac{\mathit{rules\_per\_number\_of\_contents[len(rule)]}}{\mathit{max(rules\_per\_number\_of\_contents)}}\]

Note that the only difference compared to the formula to compute
contribution of size is the distribution used, which is given by the
term $\mathit{rules\_per\_number\_of\_contents}$ denoting a dictionary
of number of content options to number of rules that contain that
number of content options.

\subsubsection{Content rarity.} 
This heuristic weighs higher rules
that use popular strings in content options. More precisely, for a
given rule \tname{} takes the arithmetic mean of the popularity scores
of each content in that rule. More precisely, we calculate the
contribution of content rarity with the following formula:

@@

\noindent
, where $N$ is the number of content options in a given rule. Figure~\Fix{...elaborate...}

@@

\begin{figure}[H]
  \centering
  \scalebox{1.2}{  
    \begin{tikzpicture}
      \begin{axis}[
          bar width=6pt,
          scaled ticks=false,
          tick label style={/pgf/number format/fixed},
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          symbolic x coords={GET, POST, SELECT, UNION, id=, DELETE, UPDATE, INSERT, 200, /index.php?},
          xtick=data,
          xticklabel style={rotate=45},          
          nodes near coords,
          every node near coord/.append style={font=\tiny},
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(GET,2007) (POST,1570) (SELECT,747) (UNION,461) (id=,408) (DELETE,390) (UPDATE,387) (INSERT, 386) (200, 352) (/index.php?, 273)};
      \end{axis}
    \end{tikzpicture}
  }
  \caption{\label{fig:distribution-contents}-}
\end{figure}

>>>>

\tname{} uses a composition of heurisitc functions to sort the rules
it produces. Each function uses a different criterion to estimate how
similar a rule, created in the previous step, is to rules from a
public database~\cite{emerging-threats-open} including a total of
\numrulessuri\ rules. We describe in the following each of these
heuristic functions. The codomain of each function ranges in the 0-1
interval. The score of a rule is the average across the values
produced by each function.

%% \Luc{atualmente eh uma
%%   média simples e não ponderada ja q as funções tem o mesmo
%%   peso. conversamos na ultima reuniao sobre testar diferentes pesos
%%   para cada função, mas n sei quão prioritário eh isso comparado as
%%   outras atividades que definimos.} $\sum{f_i}/|f_i|$.

\subsubsection{Size of a rule}
One element that \tname{} uses to measure similarity between a
synthesized and a human-created rule is the size of the synthetic
rule, \ie{}, the number of options it contains. Informally, rules with
popular sizes score high whereas inpopular rules score
low. Figure~\ref{fig:distribution-rule-size} shows the histogram of
rule sizes for our rule dataset. The x-axis shows the number of
options included in a rule (up to 10) while the y-axis shows the
number of rules with the corresponding number of options. Note that
there are 11 rules with no options. These rules only use the header to
signal a traffic as malicious. Also note that most rules have 2 to 6
options. \tname{} values rules in that size range higher compared to
other sizes.

@@

\noindent
, where $\mathit{len(r)}$ denotes the number of options of a rule $r$,
$\mathit{rules\_per\_size}$ encodes distribution of sizes as a
dictionary mapping sizes to number of associated rules, and
$\mathit{max(rules\_per\_size)}$ returns the maximum number of rules
for any dictionary entry. Consequently, the contribution of size is
proportional to its popularity. For example, the contribution of this
heuristic for a rule with three options would be 0.684 (=3,662/5,353)
whereas the contribution of a rule with nine options would be 0.028
(=154/5,353). Note that the measurement is normalized in the 0-1
range.

\subsubsection{Number of content options.} We empirically observed that content options are very
common. \Luc{Se for preciso temos numeros para fundamentar isso
  melhor, sao cerca de 70K contents nas 28K regras do
  dataset}\Mar{sim, isto ajuda, porem mais importante eh saber quantas
  (\%) das 28K regras tem contents.} As consequence, we defined a
heuristic function that takes the number of content options into
account. We calculate the contribution of number of content options
with the following formula:

@@

\[\frac{\mathit{rules\_per\_number\_of\_contents[len(r)]}}{\mathit{max(rules\_per\_number\_of\_contents)}}\]

\noindent
Note that the only difference of this formula compared to the formula
to compute contribution of size is the distribution used, which is
given by the term $\mathit{rules\_per\_number\_of\_contents}$ denoting
a dictionary of number of content options to number of rules that
contain that number of content options.

\subsubsection{Content rarity.} 
This heuristic weighs higher rules that use popular strings in content
options. For a given rule \tname{} takes the arithmetic mean of the
popularity scores of each string used in contents in that rule. More
precisely, we calculate the contribution of content rarity with the
following formula:

@@

\noindent
, where the term $\mathit{content_i}$ denotes the $i$-th xcontent in a
given rule, the term $\mathit{content\_frequency[\mathit{content_i}]}$
denotes the frequency of the string used in the $i$-th content, the
term $\mathit{max(content\_frequency)}$ denotes the maximum frequency
observed across all strings appearing as content parameters, and $N$
is the number of content options in a given
rule. Figure~\ref{fig:distribution-strings} shows the distribution of
the ten most popular strings in our rule dataset. To illustrate this
function, let us consider a rule with two content options, using the
strings ``GET'' and ``ID'', respectively. The contribution of that
rule for the computation of similarity score would be 0.601
(=(1+.203)/2).

\begin{figure}[H]
  \centering
  \scalebox{1.2}{  
    \begin{tikzpicture}
      \begin{axis}[
          bar width=6pt,
          scaled ticks=false,
          tick label style={/pgf/number format/fixed},
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          symbolic x coords={GET, POST, SELECT, UNION, id=, DELETE, UPDATE, INSERT, 200, /index.php?},
          xtick=data,
          xticklabel style={rotate=45},          
          nodes near coords,
          every node near coord/.append style={font=\tiny},
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(GET,2007) (POST,1570) (SELECT,747) (UNION,461) (id=,408) (DELETE,390) (UPDATE,387) (INSERT, 386) (200, 352) (/index.php?, 273)};
      \end{axis}
    \end{tikzpicture}
  }
  \caption{\label{fig:distribution-strings}Distribution of strings
    used as parameters of content options.}
\end{figure}

@@
